`plot.dads`

### 2023-02-24 inst/gitbook/06_plotting_dads_objects.Rmd
Update 06_plotting_dads_objects.Rmd

`"dads"` objects can be directly plotted in dads using many different options.
<!-- 
test_that("plot.dads works for traits", {

>>>>

`"dads"` objects can be directly plotted in `dads` using the S3 `plot.dads` function (or just `plot(x)` if `x` is of class `"dads"`).

# Plotting traits

`"dads"` `"traits"` objects are covered in the [traits section](#maketraits).
You can use `plot.dads` to plot them by choosing which specific trait to plot using the `trait` argument (default is 1):

```{r}
## Making a list of three traits
list_of_traits <- make.traits(process = c(no.process, BM.process, OU.process), trait.names = c("No process (normal)", "Brownian motion", "Ornstein-Uhlenbeck"))
## Plotting each trait separately
par(mfrow = c(2, 2))
plot(list_of_traits, trait = 1)
## Using different colours options
plot(list_of_traits, trait = 2, col = c("red", "purple", "pink"))
## Not using the default plot name
plot(list_of_traits, trait = 3, main = "OU process")
```

<!-- TODO: add trait labels and handling colours! -->

You can also control the number of replicates in the simulation by using the `simulations` option (the default is 50).
Bigger numbers leads to more time but smoother looking plots while smaller ones are more stochastic:

```{r}
par(mfrow = c(1,2))
plot(list_of_traits, trait = 2, simulations = 10, main = "10 BM simulations")
plot(list_of_traits, trait = 2, simulations = 1000, main = "1k BM simulations")
```

# Plotting dads results

If `dads` is used to plot only a tree (and outputs a `"phylo"` object), you can use the function `plot.phylo` from the `ape` package to plot your tree.
You'll find all the tree plotting option in the `?plot.phylo` manual page.

```{r}
## A simple pure birth tree
my_tree <- dads(stop.rule = list(max.taxa = 20))
plot(my_tree, main = "Plotting a \"phylo\" object")
```

However, if you also simulated a trait along with the tree you can use the `plot.dads` function to plot both the tree and the trait:

```{r}
## A simple pure birth tree with a BM process
my_tree <- dads(stop.rule = list(max.taxa = 20), traits = make.traits())
## Playing with the default options
plot(my_tree, main = "A tree and traits")
```

By default, elements are coloured as follows: nodes and tips are points and coloured in blue if they are tips (light blue if they are fossils) and nodes are in orange.
Branches linking them are grey lines.
You can of course change this colour palette to something of you preference by calling the normal arguments that can be passed to `points` or `lines`. For example `pch` for the point type or `lty` for the line type.
If there is an ambiguity between the argument names (for example, `cex` for both lines and points), you can specify it by prefixing the argument by the function name.
For example `cex = 2` will make both the points and lines twice bigger but `cex.points = 2` will make _only_ the points twice bigger.

```{r}
## Playing with some more options
plot(my_tree, main = "A tree and traits", points.cex = 2, pch = 19, col = c("red", "purple", "pink"), lwd = 1/3, lty = 3)
```

<!-- 
test_that("plot.dads works for traits", {
